---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
if (!file.exists("activity.csv")) { 
        unzip("activity.zip")}
dat <- read.csv("activity.csv", )
```

## What is mean total number of steps taken per day?
Total steps per day:
```{r}
totalstepsperday <- tapply(dat$steps, dat$date, sum, na.rm = TRUE)
totalstepsperday
```

Histogram, mean, and median of total steps per day:
```{r}
hist(totalstepsperday)
mean(totalstepsperday)
median(totalstepsperday)
```

## What is the average daily activity pattern?
Time Series Plot:
```{r}
meanstepsperinterval <- tapply(dat$steps, dat$interval, mean, na.rm = TRUE)
plot(names(meanstepsperinterval),
     meanstepsperinterval, 
     type="l",
     ylab="Average Number of Steps",
     xlab="5-min Interval")
```

Interval with max number of steps on average across all days:
```{r}
names(which.max(meanstepsperinterval))
```

## Imputing missing values
Total number of NAs:
```{r}
sum(is.na(dat))
```

New dataset imputing NAs with means for that 5min interval across all days:
```{r}
impdat <- transform(dat, 
                    steps = ifelse (is.na(dat$steps), meanstepsperinterval, dat$steps))
```

New histogram, mean, and median of total steps per day:
```{r}
imptotalstepsperday <- tapply(impdat$steps, impdat$date, sum)
hist(imptotalstepsperday)
mean(imptotalstepsperday)
median(imptotalstepsperday)
```

Difference in total daily number of steps, mean, and median compared to previous:
```{r}
imptotalstepsperday - totalstepsperday
mean(imptotalstepsperday) - mean(totalstepsperday)
median(imptotalstepsperday) - median(totalstepsperday)
```

## Are there differences in activity patterns between weekdays and weekends?
New factor variable with weekday/weekend.
```{r}
library(dplyr)
Day <- weekdays(as.Date(impdat$date))
impdat <- mutate(impdat, Weekendstat = factor(Day %in% c("Saturday", "Sunday"), 
                                              labels = c("Weekday", "Weekend")))
```

Panel plot:
```{r}
impmeanstepsperinterval <- aggregate(steps ~ interval + Weekendstat, impdat, mean)
library(ggplot2)
ggplot(data = impmeanstepsperinterval, aes(x = interval, y = steps)) + 
  geom_line() +
  facet_grid(Weekendstat ~ .) +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5min Interval") +
  ylab("Average Steps")
```

